Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm

### 2019-09-24 paper/virtual_cytoplasm_v2.tex
Discussion is done.

\author{Leandro Oliveira Bortot}
\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeir{\~a}o Preto, University of S{\~a}o Paulo, Ribeir{\~a}o Preto, Brazil}

\author{Zahedeh Bashardanesh}
\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}

@@

The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion} A). The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules. We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement. Thus, tRNA is considered as an outlier in our cytoplasm model. In fact, by removing diffusion coefficient of tRNA from the data, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$, is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$). These results are in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E. coli}, as measured by fluorescence recovery after photobleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.

It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}. Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{Yu2016,Mcguffee2010}. In this work we used, for the first time, the AMBER family of force fields to perform all-atom explicit solvent of a cytoplasm model for the total of \SI{3}{\micro\second}.

Except for some local unfolding for TufA, all crowders were stable during the simulations (Figure \ref{fig:structural_integrity_chain}) and their flexibility profile doesn't change significantly (Figure \ref{fig:rmsf}) when compared to the dilute condition. This shows that our cytoplasm model is stable even in the microsecond range.
The oligomeric state of the crowders... \textbf{\color{red}{MISSING RESULTS: whole crowder structural integrity using buried oligomer area}}

However, we observed aggregation of Mg, ATP and FBP...

@@

>>>>

\author{Leandro Oliveira Bortot}
\affiliation{Laboratory of Biological Physics, School of Pharmaceutical Sciences of Ribeir{\~a}o Preto, University of S{\~a}o Paulo, Ribeir{\~a}o Preto, Brazil}
\altaffiliation{These authors contributed equally to this work}

\author{Zahedeh Bashardanesh}
\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
\altaffiliation{These authors contributed equally to this work}

@@

The main difference between the behavior of the macromolecules that compose our model in the crowded system and in the dilute condition is their translational diffusion coefficient, which drops by more around 85\% for the proteins and 93\% for the tRNA (Fig.~\ref{fig:translational_diffusion}A). The slope of the line generated by fitting a linear regression model to values of $D_{cyto}/D_{dil}$ is close to zero (slope $= 0.00028$), showing that the drop in diffusion coefficient is independent of the size of the macromolecules. We found that the aggregation of ATP$^{-3}$ and FBP$^{-4}$ around Mg$^{2+}$, which was used as counter-ion for tRNA, leads to its exaggerated confinement. Thus, tRNA is considered as an outlier in our cytoplasm model. In fact, the slope of the linear regression line fitted to $D_{cyto}/D_{dil}$ is reduced by by 50\% and gets even closer to zero (slope $= 0.00014$) after removing the diffusion coefficient of tRNA from the data. Our data is in agreement with experimental results showing that the translational diffusion coefficient of Green Fluorescent Protein (GFP) in {\em E. coli}, as measured by fluorescence recovery after photobleaching (FRAP), decreased 90\% when compared to {\em in vitro} measurements~\cite{Elowitz1999,Konopka2006}.

It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}. Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{yu2016biomolecular,mcguffee2010diffusion}. In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model. All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}. The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.

All crowders are as compact in the cytoplasm model as they are in dilute conditions (Figure \ref{fig:structural_integrity_chain}A). Despite this overall stability, three crowders (TufA, MetE and IcdA) displayed greater deviation from their crystallographic structure in the cytoplasm model when compared to the dilute condition, indicating that there was some degree of local unfolding (Figure \ref{fig:structural_integrity_chain}B). Interestingly, the flexibility profiles calculated considering the C$\upalpha$ for each chain of the elements of our cytoplasm model show that, in the crowded condition, most of the proteins (TufA, MetE, AhpC, CspC and Ppa) are slightly more flexible than in the dilute condition (Figure \ref{fig:rmsf}). This is in agreement with simulation ~\cite{feig2011variable} and experimental data ~\cite{miklos2011protein,wang2012macromolecular} showing that crowders can destabilize the native structure of proteins due to the numerous unspecific short-lived interactions that are formed between them in a crowded environment.

The oligomeric state of the crowders... \textbf{\color{red}{MISSING RESULTS: whole crowder structural integrity using buried oligomer area}}

Magnesium cations were added to our cytoplasm model as counter-ions for tRNA and ATP. However, as the simulation progressed, FBP and ATP aggregated around Mg$^{2+}$ and, by consequence, around tRNA. After identifying this problem, we used simulations of small systems containing only the metabolites, Mg$^{2+}$ and water to test different strategies to prevent aggregation in future studies using long simulations of cytoplasm models. We found that protonating the phosphate-containing metabolites is enough to reach this goal (Figure \ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and FBPH$_4$ are not the biologically relevant forms of these molecules in physiological pH, we consider that using their protonated forms is a good compromise between parametrization complexity and practical results. The biggest impact of this approach is on long range interactions of these metabolites with other molecules due to the neutralization of their net charge. However, molecules in a crowded environment are restrained to smaller volumes and shorter inter molecular distances, which mitigates the effect of charge neutralization.

To the best of our knowledge, there is only one work in the literature which studied large scale cytoplasm models that included metabolites with all-atom explicit solvent molecular dynamics ~\cite{yu2016biomolecular}. Although they didn't report aggregation, they mentioned that the translational diffusion of highly charged phosphate-containing metabolites was "much slower" than expected. They attributed this observation to non-specific interactions between such metabolites and proteins. Maybe aggregation also contributed to their slower diffusion but it was not evident during \SI{140}{\nano\second}, which was the length of their longest simulation.

@@

In principle it is possible to develop new parameters for phosphate and Mg that doesn't lead to aggregation. However this is outside of the scope of this study and
